\mythesistitleNOTSYNPOSIS \tnote[fyi]{Synopsis title is ``Generalization of Consecutive Ones Property of Binary Matrices''}

### 2012-01-24 src/0031ded.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.

\definecolor{dim}{rgb}{0.7,0.7,0.7}
\def \dimspace {0.125in}

{%
  \fontencoding{T1}%
  \fontfamily{bch} %{put} %{bch} % {pbk} % \fontfamily{cmdh}
  % \fontfamily{pag}
  \selectfont%

  \vglue -0.5in % this makes top margin 2in
%   \vskip -5ex%
%   \hspace{5in}\parbox{1in}{\raggedright Hello world}

  \hspace{-1in}%
  \parbox[c]{4in}{%
    \large \raggedright \noindent %
    \color{dim}{%
      The process of preparing programs for the digital computer is
      especially attractive, not only because it can be economically
      and scientifically rewarding, but also because it can be an
      aesthetic experience much like composing poetry or music.%
      {%
        \small \par \hfill \emph{The Art of Computer Programming},
        D.\,E.\,K.%
      }%
    }%
  }%

@@

  \hspace{1.5in}%
  \parbox[t]{4in}{%
    \Large \noindent \raggedleft%
    \color{dim}{%
      The rhetoric and mythos of science create the comforting image
      of linear progression toward truth.%
      {%
        \small \par \hfill%
        \textbf{mythos}; \emph{New Oxford American Dictionary, 2nd
          Ed.}%
      }%
    }%
  }
  \\
  \vspace{\dimspace}

  % \hspace{1in}
  {\centering
    \parbox{6in}{%
%      \Huge 
      \fontsize{27}{1.5}\selectfont
      \centering \noindent %
%      \textbf{
      \emph{L'art c'est la solution au chaos.} \par \noindent
      \noindent (Art is the solution to chaos.)
%      }
%       {%
%         \small \par \hfill%
%         \emph{citation needed.}%
%       }%
    }%
  }%

@@

  \hspace{-0.25in}%
  \parbox{2.5in}{%
    \Large 
    \raggedright \noindent%
    \color{dim}{%
      Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
      the $n$.  $N = 1 \Rightarrow P = NP$ %
    }%
    {%
      \small \par \hfill%
      \emph{Concrete Mathematics}%
    }%
  }%
  \vspace{\dimspace}

  \hspace{1in}%
  \parbox{5in}{%
    \large \raggedleft \noindent%
    \color{dim}{%
      Anything that happens, happens.  Anything that, in happening,
      causes something else to happen, causes something else to
      happen.  Anything that, in happening, causes itself to happen
      again, happens again.  It doesn't necessarily do it in
      chronological order, though.  }%
    {%
      \small \par \hfill%
      \emph{The Salmon of Doubt}, D.\,N.\,A.%
    }%
  }%
  \vspace{\dimspace}

  \hspace{-0.50in}%
  \parbox{4in}{%
    \huge \noindent \raggedright%
    \color{dim}{%
      This is water, this is water.%
      {%
        \small \par \hfill%
        D.\,F.\,W.%
      }%
    }%
  }%

>>>>

\definecolor{dim}{rgb}{0.7,0.7,0.7}
\def \dimspace {0in} %{0.125in}
\def \dimspacemore {0.5in} %{0.125in}
\def \authspace {-0.25in}
\def \brightspace {0.375in}

{%
  \thispagestyle{empty}
  \fontencoding{T1}%
  \fontfamily{bch} %{put} %{bch} % {pbk} % \fontfamily{cmdh}
  % \fontfamily{pag}
  \selectfont%

  \vglue -1in % this changes top margin 
%   \vskip -5ex%
%   \hspace{5in}\parbox{1in}{\raggedright Hello world}

  \hspace{-0.75in}%
  \parbox[t]{6.5in}{%
    \Large \noindent \raggedright%
    \color{dim}{%
      % One teacher called these moments 'mathematical
      % experiences'. What I didn't know then was that
      [...] a mathematical experience was aesthetic in nature, an
      epiphany in Joyce's original sense. These moments appeared in
      proof-completions, or maybe algorithms. Or like a gorgeously
      simple solution you suddenly see after filling half a notebook
      with gnarly attempted solutions. It was really an experience of
      what I think Yeats called ``the click of a well made box''.%
      % The word I always think of is as 'click'.
      {%
        \vspace{\authspace}
        \small \par \hfill%
        D.\,F.\,W.%
      }%
    }%
  }%

@@

  \hspace{1in}%
  \parbox{4in}{ }{%
    \LARGE 
    \raggedleft \noindent%
    \color{dim}{%
      Math Graffiti: Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
      the $n$.  $N = 1 \Rightarrow P = NP$ %
      {%
%        \vspace{\authspace}%
        \small \par \hfill%
        \emph{Concrete Mathematics (margin notes)}%
      }%
    }%
  }%

@@

  \hspace{-1.25in}%
%  \ding{125}
  \parbox[t]{5in}{%
    \large \raggedright \noindent %
    \color{dim}{%
      The process of preparing programs for the digital computer is
      especially attractive, not only because it can be economically
      and scientifically rewarding, but also because it can be an
      aesthetic experience much like composing poetry or music.%
      {%
%        \vspace{\authspace}%
        \small \par \hfill \emph{The Art of Computer Programming},
        D.\,E.\,K.%
      }%
    }%
  }%

  \vspace{\brightspace}

  {%
      \centering%
      \hspace{-1.25in}
      \parbox{8in}{%
      \fontsize{36}{1.5}\selectfont%
      \centering \noindent%
      \emph{L'art c'est la solution au chaos.}%
      \par \noindent%
      \normalsize (Art is the solution to chaos.)%
    }%
  }%

  \vspace{\brightspace}

 \hspace{-0.25in}%
  \parbox[t]{6.5in}{%
    \LARGE  \noindent \raggedleft%
    \color{dim}{%
      \textbf{myth\textperiodcentered os} {|{\huge $'$mi $\theta$
          $\overline{o}$s}|} \\
      a set of beliefs or assumptions about something
      \emph{ :\\ 
      the rhetoric and mythos of science create the comforting image
      of linear progression toward truth.}%
      {%
%        \vspace{\authspace}%
        \small \par \hfill%
        \emph{New Oxford American Dictionary, 2nd
          Ed.}%
     }%
    }%
  }

  \hspace{-1in}%
  \parbox{6in}{%
    \singlespacing
    \large \raggedright \noindent%
    \color{dim}{%
      Anything that happens, happens.  Anything that, in happening,
      causes something else to happen, causes something else to
      happen.  Anything that, in happening, causes itself to happen
      again, happens again.  It doesn't necessarily do it in
      chronological order, though.%
      {%
%       \vspace{\authspace}%
        \small \par \hfill%
        \emph{The Salmon of Doubt}, D.\,N.\,A.%
      }%
    }%
  }%

  \hspace{1.5in}%
  \parbox{4in}{%
    %\huge 
    \Large 
    \noindent \raggedleft%
    \color{dim}{%
      This is water, this is water.%
      {%
%        \vspace{\authspace}%
        \small \par \hfill%
        D.\,F.\,W.%
      }%
    }%
  }%
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx
  % xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx
  % xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx xxxx xx xxx

  \vskip -02.5in % this changes bottom margin

### 2012-01-24 src/010chap1.tex
Did some over all clean up. Added fancy chapter headings. Begun work on opt survey.

\section{Organization of the document}
\label{sec:orgofdoc}
This chapter (Chapter~\ref{ch:intro}) introduces the area of research
and the problems addressed in this thesis.  Section~\ref{sec:problem}
introduces the main problem of this thesis by way of an
illustration. Section~\ref{sec:basicprelim} lays out a few general
definitions that are helpful in understanding the rest of the chapter.
Section~\ref{sec:background} gives a brief survey of COP
and optimization problems related to it followed by motivation for the
thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
presents a summary of our results on the extension of COP namely, the
tree path labeling problem.

Chapter~\ref{ch:copsurvey} gives a more detailed survey briefed in
Section~\ref{sec:background}.  Chapter~\ref{ch:myresearch} details all
the results obtained to the problems of this thesis and finally the
conclusion of the thesis is discussed in Chapter~\ref{ch:conclusion}.

@@

A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
part of their semester thesis, they pick a body of work to study and
form the namesake study groups, {\LLL}, {\GGG}, {\BBB} and
{\TTT}\tnote[minor]{put bib entries for these works!}. A student will be
in at least one study group and may be in more than one. For instance,
as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
studies only \BBB.

@@

\begin{table}[t] %[htbp]
  \centering
  {\footnotesize    
    \begin{tabular}{l||l}
      \begin{tabular}{rcl}
        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
        $n $&$=$&$ |V| = 11$\\
        $m $&$=$&$ |\cP| = 4$\\\\
        $\cl $&$=$& {\em Study group to route mapping}\\        
        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
        \cF$
      \end{tabular} & 
      \begin{tabular}{c}
        {\em Apartment allocation }($\phi$)\\
        \begin{tabular}{c|c}                      
          1  & \xSa\\
          2  & \xPi\\
          3  & \xFr\\
          4  & \xWo\\
          5  & \xCh\\
          6  & \xVi\\
          7  & \xPa\\
          8  & \xSn\\
          9  & \xLu\\
          10 & \xLi\\
          11 & \xSc                     
        \end{tabular}
      \end{tabular}\\
    \end{tabular}
  }
  \caption{\figtabsize A solution to \illustrationproblem} %study group accommodation problem
  \label{tab:iltree}
\end{table}

@@

%D<debug> figures inside tables
% \begin{tabular}{cc}
%  \begin{figure} %[t]%
% %    \centering
%     \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
% %    \label{fig:streetmap}
%     \caption{--placeholder--}
%   \end{figure}
%     &
%   \begin{figure} %[t]%
% %    \centering
%     \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
% %    \label{fig:streetmappath}
%     \caption{--placeholder--}
%   \end{figure}
% \end{tabular}

\begin{figure}[htbp] %[t]%
  \centering
  \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
  \caption{\figtabsize {\residenceblock} street map.}
  \label{fig:streetmap}

  \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
  \caption[\figtabsize \residenceblock street map with study group
    routes allocated.]{\figtabsize {\residenceblock} street map with study group
    routes allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
  \label{fig:streetmappath}
%   \caption{\figtabsize (a) {\residenceblock} street map (b)
%     {\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
%  \label{fig:streetmap}

  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
  \caption[\figtabsize Solution to the student accommodation
  problem.]{\figtabsize Individual allocation of apartments to
    students in {\residenceblock} that meets the requirements stated
    before.  The routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{yellow}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group.\\ {\tiny {\em Peanuts images
        {\copyright} Charles Schulz}}}% YellowOrange
  \label{fig:streetmappathpeople}
\end{figure} \tnote{[i] UPDATE IMAGE (b) 2-infinite-loop-BTWF.png. REMOVE
  TITLES. ADD {$\xLLL$} {$\xGGG$} {$\xBBB$} {$\xTTT$} [ii]make
  dashed/textured lines for routes. make it color agnostic.}

@@

The problem of interest in this thesis, namely, tree path labeling
problem, is a natural generalization of the interval assignment
problem or the COT problem. The problem is defined as follows -- given
a set system $\cF$ from a universe $U$ and a target tree $T$, does
there exist a bijection from $U$ to the vertices of $T$ such that each
set in the system maps to a path in $T$.  We refer to this as the
{\CFTPL} problem or simply {\em tree path labeling} problem for an
input set system and target tree pair -- $(\cF,T)$. The special case
of the target tree being a path, is the interval assignment problem.
We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
characterize feasible path assignments.  We show that for a given set
system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
sets, there is a feasible\endnote{The notion of {\em feasibility} is
  formally defined in Section~\ref{ch:prelims}.}  bijection between
$U$ and $V(T)$ if and only if the intersection cardinalities among any
three sets (not necessarily distinct) is equal to that of the
corresponding paths assigned to them and the input passes a filtering
algorithm (described in this paper) successfully.  This algorithmic
characterization gives a natural data structure that stores all the
\remove{relevant} feasible bijections between $U$ and $V(T)$. This
reduces the search space for the solution considerably from the
universe of all possible bijections between $U$ and $V(T)$ to only
those bijections that maintain the characterization.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling\endnote{The terms {\em tree path labeling} and {\em
    tree path assignment} are, in informal language,
  synonyms. Formally, the former refers to the bijection $\cl: \cF
  \rightarrow \cP$. The latter refers to the set of ordered pairs
  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
  $T$.} is feasible.

@@

We see in Section~\ref{sec:motive} that pairwise intersection
cardinality preservation is necessary and sufficient for an interval
assignment to be feasible for a given hypergraph\endnote{A {\em
    hypergraph} is an alternate representation of a set system and
  will be used in this thesis.}~\endnote{See Section~\ref{ch:prelims} for the
  formal definition.} and thus is a characterization for COP
\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
  theorems and/or lemma?} In our work we extend this characterization
and find that trio-wise intersection cardinality preservation makes a
tree path labeling\endnote{A {\em tree path labeling} $\cl$ is a
  bijection of paths from the target tree $T$ to the hyperedges in
  given hypergraph $\cF$.}~\endnotemark[4]  (TPL) feasible, which is a generalization of the COP
problem. This problem is defined as follows.

@@

We give a necessary and sufficient condition by way of {\em
  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
filtering algorithm for {\FTPL} to output in affirmative. ICPPL
captures the trio-wise cardinality property described
earlier\endnote{See Section~\ref{sec:feasible} for the definition of
  ICPPL.}. This characterization can be checked in polynomial time.  A
relevant consequence of this constructive procedure is that it is
sufficient to iteratively check if three-way intersection
cardinalities are preserved.  In other words, in each iteration, it is
sufficient to check if the intersection of any three hyperedges is of
the same cardinality as the intersection of the corresponding paths.
Thus this generalizes the well studied question of the feasible
interval assignment problem which is the special case when the target
tree $T$ is simply a path \cite{wlh02,nsnrs09}.

@@

\tnote{The following sentence seems out of place in this para} In
spite of this being a restricted case, we believe that our results are
of significant interest in understanding the nature of {\sc Graph
  Isomorphism} which is polynomial time solvable in interval graphs
while being hard on path graphs\cite{kklv10}. {\kstars} are a class
of trees which are in many ways very close to intervals or paths. Each
ray\endnote{The path from a leaf to the root, the vertex with highest
  degree, is called a {\em ray} of the \kstar.}~\endnotemark[4] are independent except for the
root\endnote{The vertex with maximum degree in a {\kstar} is called
  {\em root}.}~\endnotemark[4] and hence can be
considered as an independent interval till the root. Our algorithm
builds on this fact and uses the interval assignment
algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
the trio-wise intersection cardinality (the extra condition in ICPPL
that generalizes ICPIA) check to resolve the ambiguity about which ray
the algorithm should ``grow'' the solution into in the next iteration.

We also have an algorithm for solving {\CFTPL} with no restrictions on
the target tree or set size which runs in exponential time.  This
algorithm finds a path labeling from $T$ by decomposing the problem
into subproblems of finding path labeling of subsets of $\cF$ from
subtrees of $T$. Given the fact that binary matrices naturally
represent a set system (see Section~\ref{sec:motive}) and that the
{\em overlap}\tnote{Give informal definition of ``overlap''
  in footnote/endnote and link to sec:prelims. ADD the definition in prelim.}
relation between the sets involved is an obvious equivalence relation,
$\cF$ quite naturally partitions into equivalence classes known as
{\em overlap components}\tnote{Give informal definition of ``overlap
  components'' in footnote and link to sec:prelims. ADD the definition
  in prelim.}. In the context of COP, overlap components were used in
\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
that these equivalence classes form a total order \tnote{Give informal
  definition of ``total order'' in footnote?}. We extend this to TPL
and find that when $\cF$ is a path hypergraph\endnote{If there exists
  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
the classes can be partially ordered as an in-tree in polynomial
time. Once $\cF$ is ``broken'' into overlap components, one must
identify the subtree of $T$ that it needs to map to and this is the
hard part which is currently open to be solved in polynomial time.

@@

\theendnotes

>>>>

\section{Organization of the document}
\label{sec:orgofdoc}
Chapter~\ref{ch:intro} introduces the area of research and the
problems addressed in this thesis.  Chapter~\ref{ch:copsurvey} gives a
more detailed survey briefed in Section~\ref{sec:background}.
Chapter~\ref{ch:myresearch} details all the results obtained to the
problems of this thesis and finally the conclusion of the thesis is
discussed in Chapter~\ref{ch:conclusion}.

In this chapter, Section~\ref{sec:problem} introduces the main problem
of this thesis by way of an
illustration. Section~\ref{sec:basicprelim} lays out a few general
definitions that are helpful in understanding the rest of the chapter.
Section~\ref{sec:background} gives a brief survey of COP and
optimization problems related to it followed by motivation for the
thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
presents a summary of our results on the extension of COP namely, the
tree path labeling problem.

@@

A group of students, \Pa, \Pig, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
\Sc\ and \Lu\ enroll at the {\WSI} for a liberal arts programme.  As
part of their semester thesis, they pick a body of work to study and
form the namesake study groups, {\LLL} \cite{wallace99}, {\GGG} \cite{wallace96}, {\BBB}
\cite{wallace00} and
{\TTT} \cite{wallace10}. A student will be
in at least one study group and may be in more than one. For instance,
as will be seen later, {\Fr} studies both {\LLL} and {\TTT} while \Wo\
studies only \BBB.

@@

\begin{table}[b] %[htbp]
  \centering
  {\footnotesize    
    \begin{tabular}{l||l}
      \begin{tabular}{rcl}
        $T $&$=$&$ \text{\em Street map tree of \residenceblock}$\\
        $V(T) $&$=$&$ \{ 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11 \}$\\
        $\cP $&$=$&$ \{R\xLLL, R\xGGG, R\xBBB, R\xTTT\}$\\
        $R\xLLL $&$=$&$ \{9, 1, 5, 3, 11\}$\\
        $R\xGGG $&$=$&$ \{7, 2, 6, 5\}$\\
        $R\xBBB $&$=$&$ \{8, 2, 4 \}$\\
        $R\xTTT $&$=$&$ \{10, 6, 5, 3\}$\\
        $n $&$=$&$ |V| = 11$\\
        $m $&$=$&$ |\cP| = 4$\\\\
        $\cl $&$=$& {\em Study group to route mapping}\\        
        $\cl(\mathbb{X}) $&$=$&  $R\mathbb{X}$ for all $\mathbb{X} \in
        \cF$
      \end{tabular} & 
      \begin{tabular}{c}
        {\em Apartment allocation }($\phi$)\\
        \begin{tabular}{c|c}                      
          1  & \xSa\\
          2  & \xPi\\
          3  & \xFr\\
          4  & \xWo\\
          5  & \xCh\\
          6  & \xVi\\
          7  & \xPa\\
          8  & \xSn\\
          9  & \xLu\\
          10 & \xLi\\
          11 & \xSc                     
        \end{tabular}
      \end{tabular}\\
    \end{tabular}
  }
  \caption{\figtabsize A solution to \illustrationproblem} %study group accommodation problem
  \label{tab:iltree}
\end{table}

@@

% <debug> figures inside tables -- doesn't work. see git history to
% know what was tried.

\begin{figure}[htbp] %[t]%
  \centering
  \vspace{-10mm}
  \includegraphics[scale=0.45]{../img/1_infinite_loop__cropped.pdf}
  \caption{\figtabsize {\residenceblock} street map.}
  \label{fig:streetmap}

  \includegraphics[scale=0.45]{../img/2_infinite_loop_BTWF__cropped.pdf}
  \caption[\figtabsize \residenceblock street map with study group
    routes allocated.]{\figtabsize {\residenceblock} street map with study group
    routes allocated. % Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group
  }
  \label{fig:streetmappath}

  \includegraphics[scale=0.5]{../img/3_infinite_loop.pdf}
  \caption[\figtabsize Solution to the student accommodation
  problem.]{\figtabsize Individual allocation of apartments to
    students in {\residenceblock} that meets the requirements stated
    before.  % The routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{yellow}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group.
    \\ {\tiny {\em Peanuts images
        {\copyright} Charles Schulz}}}% YellowOrange
  \label{fig:streetmappathpeople}
\end{figure}

@@

The problem of interest in this thesis, namely, tree path labeling
problem, is a natural generalization of the interval assignment
problem or the COT problem. The problem is defined as follows -- given
a set system $\cF$ from a universe $U$ and a target tree $T$, does
there exist a bijection from $U$ to the vertices of $T$ such that each
set in the system maps to a path in $T$.  We refer to this as the
{\CFTPL} problem or simply {\em tree path labeling} problem for an
input set system and target tree pair -- $(\cF,T)$. The special case
of the target tree being a path, is the interval assignment problem.
We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
characterize feasible path assignments.  We show that for a given set
system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
sets, there is a feasible\footnote{The notion of {\em feasibility} is
  formally defined in Section~\ref{ch:prelims}.}  bijection between
$U$ and $V(T)$ if and only if the intersection cardinalities among any
three sets (not necessarily distinct) is equal to that of the
corresponding paths assigned to them and the input passes a filtering
algorithm (described in this paper) successfully.  This algorithmic
characterization gives a natural data structure that stores all the
\remove{relevant} feasible bijections between $U$ and $V(T)$. This
reduces the search space for the solution considerably from the
universe of all possible bijections between $U$ and $V(T)$ to only
those bijections that maintain the characterization.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling\footnote{The terms {\em tree path labeling} and {\em
    tree path assignment} are, in informal language,
  synonyms. Formally, the former refers to the bijection $\cl: \cF
  \rightarrow \cP$. The latter refers to the set of ordered pairs
  $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a set of paths on
  $T$.} is feasible.

@@

We see in Section~\ref{sec:motive} that pairwise intersection
cardinality preservation is necessary and sufficient for an interval
assignment to be feasible for a given hypergraph\footnote{A {\em
    hypergraph} is an alternate representation of a set system and
  will be used in this thesis.}~\footnote{See Section~\ref{ch:prelims} for the
  formal definition.} and thus is a characterization for COP
\cite{nsnrs09}. \tnote{Refer to the survey section that will list the
  theorems and/or lemma?} In our work we extend this characterization
and find that trio-wise intersection cardinality preservation makes a
tree path labeling\footnote{A {\em tree path labeling} $\cl$ is a
  bijection of paths from the target tree $T$ to the hyperedges in
  given hypergraph $\cF$.}~\footnotemark[4]  (TPL) feasible, which is a generalization of the COP
problem. This problem is defined as follows.

@@

We give a necessary and sufficient condition by way of {\em
  Intersection Cardinality Preservation Path Labeling} (ICPPL) and a
filtering algorithm for {\FTPL} to output in affirmative. ICPPL
captures the trio-wise cardinality property described
earlier\footnote{See Section~\ref{sec:feasible} for the definition of
  ICPPL.}. This characterization can be checked in polynomial time.  A
relevant consequence of this constructive procedure is that it is
sufficient to iteratively check if three-way intersection
cardinalities are preserved.  In other words, in each iteration, it is
sufficient to check if the intersection of any three hyperedges is of
the same cardinality as the intersection of the corresponding paths.
Thus this generalizes the well studied question of the feasible
interval assignment problem which is the special case when the target
tree $T$ is simply a path \cite{wlh02,nsnrs09}.

@@

\tnote{The following sentence seems out of place in this para} In
spite of this being a restricted case, we believe that our results are
of significant interest in understanding the nature of {\sc Graph
  Isomorphism} which is polynomial time solvable in interval graphs
while being hard on path graphs\cite{kklv10}. {\kstars} are a class
of trees which are in many ways very close to intervals or paths. Each
ray\footnote{The path from a leaf to the root, the vertex with highest
  degree, is called a {\em ray} of the \kstar.}~\footnotemark[4] are independent except for the
root\footnote{The vertex with maximum degree in a {\kstar} is called
  {\em root}.}~\footnotemark[4] and hence can be
considered as an independent interval till the root. Our algorithm
builds on this fact and uses the interval assignment
algorithm\cite{nsnrs09} up until ``reaching'' the root and then uses
the trio-wise intersection cardinality (the extra condition in ICPPL
that generalizes ICPIA) check to resolve the ambiguity about which ray
the algorithm should ``grow'' the solution into in the next iteration.

We also have an algorithm for solving {\CFTPL} with no restrictions on
the target tree or set size which runs in exponential time.  This
algorithm finds a path labeling from $T$ by decomposing the problem
into subproblems of finding path labeling of subsets of $\cF$ from
subtrees of $T$. Given the fact that binary matrices naturally
represent a set system (see Section~\ref{sec:motive}) and that the
{\em overlap}\tnote{Give informal definition of ``overlap''
  in footnote/endnote and link to sec:prelims. ADD the definition in prelim.}
relation between the sets involved is an obvious equivalence relation,
$\cF$ quite naturally partitions into equivalence classes known as
{\em overlap components}\tnote{Give informal definition of ``overlap
  components'' in footnote and link to sec:prelims. ADD the definition
  in prelim.}. In the context of COP, overlap components were used in
\cite{wlh02} and \cite{kklv10}. Moreover, \cite{nsnrs09} discovered
that these equivalence classes form a total order \tnote{Give informal
  definition of ``total order'' in footnote?}. We extend this to TPL
and find that when $\cF$ is a path hypergraph\footnote{If there exists
  an FTPL for a hypergraph $\cF$, it is called a path hypergraph.},
the classes can be partially ordered as an in-tree in polynomial
time. Once $\cF$ is ``broken'' into overlap components, one must
identify the subtree of $T$ that it needs to map to and this is the
hard part which is currently open to be solved in polynomial time.

@@

%\theendnotes
